---
title: "Tutorials Points R Language"
output: html_notebook
---

**Resources From**: https://www.tutorialspoint.com/r/r_excel_files.htm

**Data Types in R**

1. Logical *e.g. TRUE, FALSE*

2. Numeric (Floating point)

3. Integer 

4. Character - by double or single quotes

5. Raw - ``charToRaw("string")``

```{r}
x <- "Someting";
y <- charToRaw("Someone")

print(x)
print(y)
print(class(y))
```

**Vectors**
```{r}
# vector only one data types
vector1 = c("stuA", "stuB", "stuC")
print(vector1)

# list has multiple data types
list1 <- list(c(1,2,3), sin, "someone")
print(list1)

# Matrices (as two dimensions matrix)
M <- matrix( c('a','a','b','c','b','a'), nrow = 2, ncol = 3, byrow = FALSE)
print(M)

# Arrays as multidimention arrays
array1 <- array(c('green','yellow'),dim = c(3,3,2))
print(array1)

# Factors for vectors
apple_colors <- c('green','green','yellow','red','red','red','green')

# Create a factor object.
factor_apple <- factor(apple_colors)

# Print the factor.
print(factor_apple)
print(nlevels(factor_apple))

# Dataframes
BMI <- 	data.frame(
   gender = c("Male", "Male","Female"), 
   height = c(152, 171.5, 165), 
   weight = c(81,93, 78),
   Age = c(42,38,26)
)
print(BMI)

```

**Variables**

- Has letters, numbers, dot and underscore

- dot can be as start but cannot be followed by a numbers

*Unusual Assignment*
``variables -> variablename``

```{r}
# String Concatenation

# cat("string1", var, "\n")

# finding the list of variables in the memories
print(ls())

# remove a variable
rm(varname)

```

**Operators**

1. Addition

2. Subtraction

3. Multiplication

4. Division

5. ``%%`` is the remainders

6. ``%/%`` is the quotients

7. ``^`` means power

- *most of the operators in python* are applicable here

**Miscellaneous**

```{r}

# ":" means "until"
v <- 20:30
print(v)

# %in%
v1 <- 8;
v2 <- 1:10
print(v1 %in% v2)

# %*% is used to multiply a maxtrix with its transpose
M = matrix( c(2,6,5,1,10,4), nrow = 2,ncol = 3,byrow = TRUE)
t = M %*% t(M)
print(t)
```

**Control Flow**

**If-statement**

```{r}
# if(boolean_expression) {
#    // statement(s) will execute if the boolean expression is true.
# }
x <- 2;
if(x == 3) {
   print("Statement is evaluated to be true. ")
} else {
  print("Statement is evaluated to be false. ")
}
```

**If-Else Stament (Nested)**

```{r}
# if(boolean_expression) {
#    // statement(s) will execute if the boolean expression is true.
# } else {
#    // statement(s) will execute if the boolean expression is false.
# }

x <- 5;
if(x == 3) {
  print("X is equivalent to 3.")
} else if (x==4) {
  print("X is equivalent to 4. ")
} else {
  print("X is not equivalent to 3 or 4.")
}

```

**Switch**

```{r}
x <- switch(
   3,
   "first",
   "second",
   "third",
   "fourth"
)
print(x)
```

**Looping in R**

**Repeat**
```{r}
# repeat { 
#    commands 
#    if(condition) {
#       break
#    }
# }
x <-  1;

repeat {
  print(x)
  x <- x + 1;
  if (x >= 5) {
    cat(x, " is now five, terminating")
    break
  }
}

```

**While Loops**
```{r}
# while (test_expression) {
#    statement
# }

v <- c("Hello","while loop")
cnt <- 2

while (cnt < 7) {
   print(v)
   cnt = cnt + 1
}

```

**For Loops**
```{r}
# for (value in vector) {
#    statements
# }
v <- LETTERS[1:4]
for ( i in v) {
   print(i)
}

# %in% is used to identify if an element belongs to a vector.
```

**Next Statement**
```{r}
v <- LETTERS[1:6]
for ( i in v) {
   
   if (i == "D") {
      next
   }
   print(i)
}
```

**Users-Defined Functions**
```{r}
# defining a function
mysquare <- function(vectorinput, vectorinput2=1:3) {
  for (i in vectorinput) {
    cat(sqrt(i), " is the squared-root value.\n")
  }
  
  for (j in vectorinput2) {
    cat(j**2, " is the squared value.\n")
  }
  
  return("Execution Finished")
}

# invoking a function
x <- c(1,2,3,5)
y = c(4,4,5)
# mysquare(vectorinput2 = x, vectorinput = y)
funcout = mysquare(x)
print(funcout)
```

**Working with String**
```{r}
a <- "Hellow"
b <- "my name is"
c <- "nobody"

print(paste(a, b,c, sep="-", collapse = ""))
# collapse is used to eliminate the space in between two strings. 

```

**Text-Formatting**
```{r}
# format(x, digits, nsmall, scientific, width, justify = c("left", "right", "centre", "none")) 

res <- format(23.123456789, digits = 4)
print(res)

# Functions in a string
mystring <- "This is a string"
print(nchar(mystring))
print(toupper(mystring))
print(tolower(mystring))

# substring(x,first,last)
print(substring(mystring, 2, 4))
```

**Working with vectors**
```{r}
print(seq(1, 11, by=2))

# Accessing an arrays
array1 <- c(1, "Text", TRUE, 3.5)
print(array1[4])

# vector arithmetic operation is possible and broadcating is applicable in R-language also

# Sorting an array
vect1 = v(1,6,2,7,3,9,5)
print(sort(vect1, decreasing = TRUE))

```

**Working with List**
```{r}
# Naming the list
listdata <- list(c("Jan", "Feb", "Mac"), c("studentA", "StudentB"), c(157, 187, 190))
names(listdata) = c("Month", "Student Name", "Height")
print(listdata)

# Assessing the element in the list using the name
print(listdata[2])
print(listdata$`Student Name`)

# converting list into vector
# unlist(listname)

```

**Working with Matrix**
```{r}
# matrix(data, nrow, ncol, byrow, dimnames)
M <- matrix(c(3:14), nrow = 4, byrow = FALSE)
print(M)
rownames = c("row1", "row2", "row3", "row4")
colnames = c("col1", "col2", "col3")
P <- matrix(c(3:14), nrow = 4, byrow = TRUE, dimnames = list(rownames, colnames))
print(P)
print(P[2,3])
print(P[, 3])
print(P[3, ])

# Matrix Arithmetic Operation is possible
```

**Working with Arrays**
```{r}
# Create two vectors of different lengths.
vector1 <- c(5,9,3)
vector2 <- c(10,11,12,13,14,15)

# Take these vectors as input to the array.
result <- array(c(vector1,vector2),dim = c(3,3,2))
print(result)

# Naming the Arrays
# Create two vectors of different lengths.
vector1 <- c(5,9,3)
vector2 <- c(10,11,12,13,14,15)
column.names <- c("COL1","COL2","COL3")
row.names <- c("ROW1","ROW2","ROW3")
matrix.names <- c("Matrix1","Matrix2")

# Take these vectors as input to the array.
result <- array(c(vector1,vector2),dim = c(3,3,2),dimnames = list(row.names,column.names,
   matrix.names))
print(result)
print(result[1,3,1])
print(result[1,,1])

# Apply function
# apply(arrayname, margin (name of the datast to be used), function)
# Create two vectors of different lengths.
vector1 <- c(5,9,3)
vector2 <- c(10,11,12,13,14,15)

# Take these vectors as input to the array.
new.array <- array(c(vector1,vector2),dim = c(3,3,2))
print(new.array)

# Use apply to calculate the sum of the rows across all the matrices.
result <- apply(new.array, c(1), sum)
print(result)

```

**Factors Function**
```{r}
# used to categorize the data and store it as levels
mydata <- c("East","West","East","North","North","East","West","West","West","East","North")

myfactor <- factor(mydata)
print(myfactor)
print(is.factor(myfactor))

# Generating Factor Levels
# gl(nintegers given in labels, k(number of replications), labels=c(nth element))
```

**Creating Data Frame**
```{r}
# Create the data frame.
emp.data <- data.frame(
   emp_id = c (1:5), 
   emp_name = c("Rick","Dan","Michelle","Ryan","Gary"),
   salary = c(623.3,515.2,611.0,729.0,843.25), 
   
   start_date = as.Date(c("2012-01-01", "2013-09-23", "2014-11-15", "2014-05-11",
      "2015-03-27")),
   stringsAsFactors = FALSE
)
# Print the data frame.			
print(emp.data) 

# Get the structure of the data frame.
str(emp.data) 

# something like summary in pandas (python)
print(summary(emp.data))  

# Extract Specific columns.
result <- data.frame(emp.data$emp_name,emp.data$salary)
print(result)

# Extract first two rows.
result <- emp.data[1:2,]

# Extract 3rd and 5th row with 2nd and 4th column.
result <- emp.data[c(3,5),c(2,4)]

# Add the "dept" column.
emp.data$dept <- c("IT","Operations","IT","HR","Finance")

# Bind the two data frames. (adding rows records numbers)
emp.finaldata <- rbind(emp.data,emp.newdata)

# Combine above vectors into one data frame.
addresses <- cbind(city,state,zipcode)

```

**Working with Packages**
```{r}
# Checking the library asseible paths
.libPaths()

# get the list of installed library
library()

# installing packages from CRAN
install.packages("package name")

# loading a library
library('package name', lib.loc='path to library', )

```

**Working With CSV Files**
```{r}
print(getwd())
# setwd("/path/")
mydata = read.csv("mydata.csv")
print(mydata)

# Analyzing CSV files
print(is.data.frame(data))
print(ncol(data))
print(nrow(data))

# Get the person detail having max salary.
retval <- subset(data, salary == max(salary))
print(retval)

# Writting a CSV file
retval <- subset(data, as.Date(start_date) > as.Date("2014-01-01"))
write.csv(retval,"output.csv")

```
















